Topography and canopy height differently shape the functional composition of canopy and understory guilds in an Indian rainforest

### 2020-10-13 Manuscript.Rmd
Uppangala 0.1.9058 13.10.20 Reboot

---
title: Different environment filters shape the functional composition of canopy and understory guilds in an Indian rainforest
date: '`r Sys.Date()`'
output:
  bookdown::html_document2:
    number_sections: false  
    toc: true
    toc_float: yes
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: false
    toc: false
  bookdown::word_document2:
    reference_docx: ./template/template.docx
linestretch: 1.5
csl: /home/sylvain/Documents/Bibliography/csl/joe.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
rm(list = ls()); invisible(gc())
library(raster)
library(parallel)
library(knitr)
library(kableExtra)
if(knitr:::is_html_output()) options(knitr.table.format = "html") 
if(knitr:::is_latex_output()) options(knitr.table.format = "latex") 
library(tidyverse)
library(broom)
library(reshape2)
library(ggfortify)
theme_set(bayesplot::theme_default())
opts_chunk$set(echo = F, message = F, warning = F, 
               fig.height = 8, fig.width = 8,
               cache = F, cache.lazy = F)
cv <- function(x, na.rm = T)
  sd(x, na.rm = na.rm)/mean(x, na.rm = na.rm)
n = 999 # Random repetitions
sapply(list.files("R", full.names = T), source)
```

```{r data}
# Sp level PFT
PFT <- genPFT() %>% 
  mutate(LA = log(LA)) %>% 
  mutate(SLA = log(SLA)) %>% 
  select(-CE) %>% 
  group_by(Tree, Sp, SpCode) %>% 
  summarise_all(funs(mean(., na.rm = T))) %>% 
  ungroup() %>% 
  select(-Tree) %>% 
  group_by(Sp, SpCode) %>% 
  summarise_all(funs(mean(., na.rm = T))) %>% 
  ungroup() %>% 
  select(-LTD, -SHA, -Sp)
# PCA
PFT.pca <- princomp(~ LA + WD + LT + LDMC + SLA, data = PFT, cor = T)
# Matrix Q
PFT <- PFT %>% 
  mutate(SpCode = as.character(SpCode)) %>% 
  left_join(genSpecies(), by = "SpCode") %>% 
  mutate(strata = ifelse(Strata == 3, "U", "C")) %>%
  mutate(strata = ifelse(SpCode == "agin", "U", strata)) %>% 
  dplyr::select(strata, SpCode, LA, WD, LT, LDMC, SLA) %>% 
  reshape2::melt(id.vars = c("strata", "SpCode"), 
                 variable.name = "trait", value.name = "traitvalue")
# Matrix L
com <- quadrats(cs = c(30,30))
XY <- genTrees()[c('x', 'y')]
coordinates(XY) <- ~ x + y
proj4string(XY) <- crs(com)
trees <- genTrees() %>% 
  mutate(community = (XY %over% com)[,1]) %>% 
  filter(!is.na(community)) %>% 
  group_by(community, SpCode) %>% 
  summarise(Abundance = n())
# Matrix R
load(file.path("draft_save", 'R.Rdata'))
# GWM
GWM <- trees %>% 
  left_join(PFT) %>% 
  filter(!is.na(strata)) %>% 
  group_by(strata, community, trait) %>% 
  summarise(GWM = weighted.mean(traitvalue, Abundance, na.rm = T))
# GWM plot
left_join(GWM, R) %>% 
  reshape2::melt(id.vars = c("strata", "community", "trait", "GWM")) %>% 
  filter(variable != "BA") %>% 
  ggplot(aes(value, GWM, col = strata)) +
  facet_grid(trait ~ variable, scales = "free") +
  geom_point() +
  geom_smooth(method = "lm") +
  theme(legend.position = "bottom", axis.title = element_blank())
# nGWM
nGWM <- lapply(1:n, function(n) 
  left_join(trees, 
            group_by(PFT, trait, strata) %>% 
              mutate(SpCode = sample(SpCode)),
            by = "SpCode") %>% 
    filter(!is.na(strata)) %>% 
    group_by(strata, community, trait) %>% 
    summarise(nGWM = weighted.mean(traitvalue, Abundance, na.rm = T)) %>% 
    mutate(NullModel = n)) %>% 
  bind_rows()
# nGWM plot
ggplot(nGWM, aes(nGWM, col = community), fill = NA) +
  geom_density(alpha = .5) +
  facet_wrap(~ strata + trait, scales = "free") +
  scale_color_discrete(guide = "none")
# SES
SES <- GWM %>% 
  left_join(group_by(nGWM, strata, community, trait) %>% 
              summarise(nGWMmean = mean(nGWM, na.rm = T),
                        nGWMsd = sd(nGWM, na.rm = T))) %>% 
  mutate(SES = (GWM - nGWMmean)/nGWMsd) %>% 
  dplyr::select(-GWM, -nGWMmean, -nGWMsd) %>% 
  left_join(R)
# SES plot
reshape2::melt(SES, id.vars = c("strata", "community", "trait", "SES")) %>% 
  filter(variable != "BA") %>% 
  ggplot(aes(value, SES, col = strata)) +
  facet_grid(trait ~ variable, scales = "free") +
  geom_point() +
  geom_hline(yintercept = c(-2, 0, 2), linetype = "dashed") +
  geom_smooth(method = "lm") +
  theme(legend.position = "bottom", axis.title = element_blank())
```

1. Sylvain Schmitt
    - Corresponding author
    - sylvain.m.schmitt@gmail.com
    - orcid.org/0000-0001-7759-7106
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
    - UMR BIOGECO, INRA - Université de Bordeaux, 69 route d’Arcachon, F-33612 Cestas cedex France
1. Valérie Raevel
    - vraevel@gmail.com
    - orcid.org/0000-0002-9631-114X
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
1. Maxime Réjou-Méchain
    - maxime.rejou@ird.fr
    - orcid.org/0000-0003-2824-267X
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
    - AMAP, IRD, CNRS, CIRAD, INRA, Univ Montpellier, Montpellier, France
1. Narayanan Ayyappan
    - ayyappan.n@ifpindia.org
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
1. Natesan Balachandran
    - balachandran.n@ifpindia.org
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
1. N. Barathan
    - barathan.n@ifpindia.org
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
1. Chandra Shekhar Jha
    - email
    - Forestry and Ecology Group, National Remote Sensing Centre, Hyderabad 500 037, India
1. François Munoz
    - fmunoz@univ-grenoble-alpes.fr
    - orcid.org/0000-0001-8776-4705
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
    - Laboratoire d'Écologie Alpine, Univ. Grenoble Alpes, FR-38000 Grenoble, France

@@

@@

1. Topographic heterogeneity determines changing soil conditions, and environment also changes across vertical strata in rainforests. However, the joint influence of these factors on tree functional composition remain little known.
<!-- Joint study of topography and community vertical structure hold promise to understand the drivers of hyperdiverse tropical communities.  -->
<!-- Moreover, understory species are generally characterized by a low potential growth rate associated to a conservative strategy of resource use, while canopy species tend to have a higher potential growth rate associated to a faster resource acquisition.  -->
<!-- Despite, assembly of tree communities has been mostly studied irrespectively of vertical stratification in tropical forests. -->
We hypothesized different leaf and wood functional strategies between canopy and understory guilds of trees, and different responses of the guilds to topographic heterogeneity, depending on resource availability, hydraulic and biomechanic constraints.
1. We measured leaf and wood traits over 89 species representing 99% of all individuals in a 10-ha rainforest plot located in the Western Ghats, South India. 
The plot was subdivided into tree communities, each composed of guilds of canopy and understory trees.
We combined tree census data, LiDAR-derived topographic and canopy metrics, and functional traits to assess the variation of functional composition in understory and canopy guilds, and to address the effects of abiotic and biotic environments on these compositions.  
1. The variation of functional traits across sampled tree species followed decoupled wood and leaf economics spectra, even at local scale. 
Although functional space did not differ overall between canopy and understory tree species, environmental filtering shaped different functional composition of canopy and understory guilds in tree communities. 
Furthermore, the functional composition of the canopy and understory guilds responded differently across communities to environmental gradients related to soil resource, light availability and aspect.
1. *Synthesis.* These results show that both fine-scale environmental heterogeneity and vertical structure shape the composition of hyperdiverse rainforest tree communities.
Guild-specific responses to biotic and abiotic environmental filtering underline that vertical stratification must be acknowledged to better understand and forecast forest dynamics.

@@

<!-- > To be checked at the end -->
Rainforest; Western Ghats; Economics spectra; Environmental filtering, Vertical strata

@@

Understanding the drivers of tree composition in hyperdiverse tropical forests is a long-standing issue in ecology [@Cody1976]. 
Environmental variation influences composition of tree community at regional [@Rejou-Mechain2008; @Tuomisto2003]  and local scale [@Ferry2007; @gunatilleke_specieshabitat_2006]. 
Specifically, topography is known to filter different ecological strategies [@kraft_functional_2008] due to its impact on water and nutrient availability [@ferry2010higher; @Blanchard2019].
<!-- At higher topographic positions, e.g., hilltops, decreased water and nutrient availability  [@ferry2010higher] is expected to filter species with more conservative resource-use strategy [@Cunningham2007; @kraft_functional_2008; @Mendez-Toribio2017; Schmitt et al., under review]. -->
However, tree species can occupy different vertical strata at adult stage, and thus be classified into vertical guilds (understory and canopy), experimenting contrasted environmental conditions. 
Indeed, light availability decreases downward across vertical forest strata, while hydraulic and mechanical constraints increase upward, so that different survival and growth strategies should be filtered across strata [@Raevel2018; @Stark2015; @Uriarte2004a].
While the influence of ground-level variation in environmental conditions (soil, topography) on community assembly is now widely recognized, environmental filtering across vertical guilds is little addressed in tropical forests [@wright2002plant].

<!-- In addition, different responses of vertical strata to environmental gradients have been observed in Columbia [@duque2002different] and central Africa [@Ouedraogo2013]. -->

Functional traits express fundamental trade-offs determining species occurrences and co-occurrences along environmental gradients [@Wright2002], contributing thus to test for community assembly processes. Functional traits have been defined as phenotypic traits that impact fitness indirectly through their effect on individual performance, which is defined as the ability to recruit, grow, survive, and reproduce in a given environment [@violle_let_2007]. 
Understory species are generally characterized by a low potential growth rate associated to a conservative strategy of resource use, while canopy species tend to have a higher potential growth rate associated to a faster resource acquisition [@Herault2011].
However, functional traits covary among species [@diaz_global_2015] and communities [@Bruelheide2018] along distinct economics spectra of leaf [@wright_worldwide_2004; @osnas_global_2013; but see @lloyd_photosynthetically_2013] and wood [@chave_towards_2009]. The leaf economics spectrum expresses a trade-off between resource acquisition by photosynthesis and investment in leaf defense and durability [@Evans2001a; @hodgson_is_2011; but see @Gibert2016]. 
Trait values along the leaf economics spectrum are expected to change along soil and light availability gradients.
The wood economics spectrum opposes slow growing, drought-resistant species to fast growing, drought-intolerant species [@chave_towards_2009].
Previous studies showed decoupled trait variations along the leaf and wood economics spectra for tropical trees at large scale in the Neotropics [@Baraloto2010; @fortunel_leaf_2012].
Nevertheless, @reich_world-wide_2014 proposed a unique plant economics spectrum simply opposing fast and slow growing species as observed in subarctic flora [@freschet_evidence_2010].

<!-- Here we decompose a tree community in spatially circumscribed at ground level into vertical guilds. Because they experience different environmental constraints, vertical guilds should show different functional responses to environmental variation between communities [@Raevel2018].  -->
In the context of seasonal rainforests, we hypothesized that greater hydraulic constraints should entail stronger conservative resource-use strategy [@Cunningham2007; @kraft_functional_2008; @Mendez-Toribio2017; Schmitt et al., under review] of canopy trees to water and nutrient limitation on hilltops [@ferry2010higher], 
 <!-- In the context of seasonal rainforests, we hypothesized that greater hydraulic constraints should entail stronger response of canopy trees to water limitation on hilltops (along the wood economics spectrum), -->
while limited light availability should favor conservative strategies in understory trees (along the leaf economics spectrum). Thus we expected different trait-environment relationships across vertical guilds, depending on tree species strategies along the fundamental leaf and wood economics spectra. Previous works have shown different response of vertical strata to environmental variation at ground level, in Columbia [@duque2002different] and central Africa [@Ouedraogo2013].

We assessed the variation in functional composition of understory and canopy guilds within tropical tree communities, and examined the effects of abiotic and biotic environments on these variations. We measured leaf and wood traits of 89 species representing 99% of all individuals in a 10-ha forest plot located in the Western Ghats, South India, a region considered as a global biodiversity hotspot [@myers_biodiversity_2000]. The plot encompasses a broad variation of topography [@Antin2013] and vertical structure [@Pascal1988].
Because the region experiences a long dry season period and is under the influence of intense monsoon during rainy seasons with strong wind disturbances, we expected hydraulic and mechanic constraints to play an important role in tree community assembly.

@@

<!-- Based on previous evidences of independent economics spectra at large scale in the Neotropics [@baraloto_decoupled_2010; @fortunel_leaf_2012], we would expect leaf and wood economics spectra to be maintained in Indian rainforests at local scale [but see @Messier2017], with differentiated trait spaces for canopy and understory species. We hypothesized functional composition to differ between canopy and understory species, due to their differences in growth and functional strategies [@Herault2011]. Because this region experiences a long dry season period and is under the influence of intense monsoon rainy seasons with strong wind disturbances, we expected that the local topography is a major driver of tree community functional composition.  -->

@@

Using the classification of @Pascal1988, we assigned a potential vertical stratum to each tree species (vertical stratum reached by most adult trees, with four categories: emergent above 39 m, canopy between 25 and 39 m, sub-canopy between 15 and 25 m, and understory below 15 m. Because our study site is characterized by a steep topography, sub-canopy species can often reach the canopy in the Uppangala plot. We thereby defined two broader categories, namely, canopy guild (C) including emergent, canopy and sub-canopy species and understory guild (U) composed of understory trees.

We collected trait data for 89 over the 111 (80%) species, representing 99% of tree individuals in the plot (54 canopy and 35 understory species). We sampled five individuals per species for the 32 most abundant species (>80% of individuals), and one to five individuals for the remaining species (3 individuals on average), depending on leaf accessibility and species abundance. For each individual, we collected at 4 to 12 m height five mature leaves and one branch sample with a diameter between 1 and 2 cm, following recommendation of @Perez-Harguindeguy2013. Whenever possible, individuals and leaves were selected in similar light conditions (Dawkins index of 2; [@Dawkins1958]). Leaf and branch samples were kept in humidified ziplock bags filled with enriched $CO_2$ air in dark until measurement within the next 24 hours. For 11 species (1% of individuals in Uppangala plot), branch samples were too small (< to 1 cm in diameter) to allow measuring branch wood density. 
<!-- Among the selected species, 54 and 35 belonged to the canopy and understory guilds, respectively. -->

@@

@@

We subsampled tree communities in $30~m*30~m$ subplots (n=110), in order to characterize the functional composition of vertical guilds within each subplot. Each subplot contained 28 individuals and 10 species on average, with less species in understory (6 in average) than in canopy (13 in average) guilds.

We thus calculated the weighted mean trait value per guild (GWM, where G stands for ‘guild’; [@Raevel2018]) for each trait in each subplot:

@@

Functional traits data have been submitted to TRY initiative [@Kattge2019] under the name UppangalaTraits. Environmental variables and spatial positions of individuals were extracted from the Uppangala Station database or **lidar data**, for which access is modulated by **to precise**.

>>>>

---
title: Topography and canopy height differently shape the functional composition of canopy and understory guilds in an Indian rainforest
date: '`r Sys.Date()`'
output:
  bookdown::word_document2:
    reference_docx: ./template/template.docx
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: false
    toc: false
  bookdown::html_document2:
    number_sections: false  
    toc: true
    toc_float: yes
linestretch: 1.5
csl: /home/sylvain/Documents/Bibliography/csl/joe.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
rm(list = ls()); invisible(gc())
library(raster)
library(parallel)
library(knitr)
library(kableExtra)
if(knitr:::is_html_output()) options(knitr.table.format = "html") 
if(knitr:::is_latex_output()) options(knitr.table.format = "latex") 
library(tidyverse)
library(broom)
library(reshape2)
library(ggfortify)
library(rstan)
options(mc.cores = 4)
rstan_options(auto_write = T)
library(bayesplot)
theme_set(bayesplot::theme_default())
opts_chunk$set(echo = F, message = F, warning = F, 
               fig.height = 8, fig.width = 8,
               cache = T, cache.lazy = F)
cv <- function(x, na.rm = T)
  sd(x, na.rm = na.rm)/mean(x, na.rm = na.rm)
n = 999 # Random repetitions
sapply(list.files("R", full.names = T), source)
```

```{r data, include = F}
# Sp level PFT
PFT <- genPFT() %>% 
  mutate(LA = log(LA)) %>% 
  mutate(SLA = log(SLA)) %>% 
  select(-CE) %>% 
  group_by(Tree, Sp, SpCode) %>% 
  summarise_all(funs(mean(., na.rm = T))) %>% 
  ungroup() %>% 
  select(-Tree) %>% 
  group_by(Sp, SpCode) %>% 
  summarise_all(funs(mean(., na.rm = T))) %>% 
  ungroup() %>% 
  select(-LTD, -SHA, -Sp)
# PCA
PFT.pca <- princomp(~ LA + WD + LT + LDMC + SLA, data = PFT, cor = T)
# Matrix Q
PFT <- PFT %>% 
  mutate(SpCode = as.character(SpCode)) %>% 
  left_join(genSpecies(), by = "SpCode") %>% 
  mutate(strata = ifelse(Strata == 3, "U", "C")) %>%
  mutate(strata = ifelse(SpCode == "agin", "U", strata)) %>% 
  dplyr::select(strata, SpCode, LA, WD, LT, LDMC, SLA) %>% 
  reshape2::melt(id.vars = c("strata", "SpCode"), 
                 variable.name = "trait", value.name = "traitvalue")
# Matrix L
com <- quadrats(cs = c(30,30))
XY <- genTrees()[c('x', 'y')]
coordinates(XY) <- ~ x + y
proj4string(XY) <- crs(com)
trees <- genTrees() %>% 
  mutate(community = (XY %over% com)[,1]) %>% 
  filter(!is.na(community)) %>% 
  group_by(community, SpCode) %>% 
  summarise(Abundance = n())
# Matrix R
load(file.path("draft_save", 'R.Rdata'))
# GWM
GWM <- trees %>% 
  left_join(PFT) %>% 
  filter(!is.na(strata)) %>% 
  group_by(strata, community, trait) %>% 
  summarise(GWM = weighted.mean(traitvalue, Abundance, na.rm = T))
# GWM plot
left_join(GWM, R) %>% 
  reshape2::melt(id.vars = c("strata", "community", "trait", "GWM")) %>% 
  filter(variable != "BA") %>% 
  ggplot(aes(value, GWM, col = strata)) +
  facet_grid(trait ~ variable, scales = "free") +
  geom_point() +
  geom_smooth(method = "lm") +
  theme(legend.position = "bottom", axis.title = element_blank())
# nGWM
nGWM <- lapply(1:n, function(n) 
  left_join(trees, 
            group_by(PFT, trait, strata) %>% 
              mutate(SpCode = sample(SpCode)),
            by = "SpCode") %>% 
    filter(!is.na(strata)) %>% 
    group_by(strata, community, trait) %>% 
    summarise(nGWM = weighted.mean(traitvalue, Abundance, na.rm = T)) %>% 
    mutate(NullModel = n)) %>% 
  bind_rows()
# nGWM plot
# ggplot(nGWM, aes(nGWM, col = community), fill = NA) +
#   geom_density(alpha = .5) +
#   facet_wrap(~ strata + trait, scales = "free") +
#   scale_color_discrete(guide = "none")
# SES
SES <- GWM %>% 
  left_join(group_by(nGWM, strata, community, trait) %>% 
              summarise(nGWMmean = mean(nGWM, na.rm = T),
                        nGWMsd = sd(nGWM, na.rm = T))) %>% 
  mutate(SES = (GWM - nGWMmean)/nGWMsd) %>% 
  dplyr::select(-GWM, -nGWMmean, -nGWMsd) %>% 
  left_join(R)
# SES plot
# reshape2::melt(SES, id.vars = c("strata", "community", "trait", "SES")) %>% 
#   filter(variable != "BA") %>% 
#   ggplot(aes(value, SES, col = strata)) +
#   facet_grid(trait ~ variable, scales = "free") +
#   geom_point() +
#   geom_hline(yintercept = c(-2, 0, 2), linetype = "dashed") +
#   geom_smooth(method = "lm") +
#   theme(legend.position = "bottom", axis.title = element_blank())
```

```{r, eval=F}
filter(trees, community %in% filter(R, Canopy < 20, Slope > 30)$community) %>% 
  group_by(community, SpCode) %>% 
  summarise(Abundance = sum(Abundance)) %>% 
  left_join(filter(PFT, trait == "WD")) %>% 
  filter(strata == "U") %>% 
  arrange(community, desc(traitvalue)) %>% 
  kable(caption = "Community with low canopy (<20m) and strong slope (>30)")
filter(trees, community %in% filter(R, Slope < 15)$community) %>% 
  group_by(community, SpCode) %>% 
  summarise(Abundance = sum(Abundance)) %>% 
  left_join(filter(PFT, trait == "WD")) %>% 
  filter(strata == "U") %>% 
  arrange(community, desc(traitvalue)) %>% 
  kable(caption = "Community with low slope (<15)")
filter(PFT, strata == "U", trait == "WD") %>% 
  ggplot(aes(traitvalue)) +
  geom_histogram() +
  ggtitle("WD distribution among understory species.")
left_join(trees, PFT) %>% 
  filter(!is.na(strata)) %>% 
  group_by(strata, community) %>% 
  summarise(Abundance = sum(Abundance)) %>% 
  left_join(R) %>% 
  ggplot(aes(Slope, Abundance)) +
  geom_point() +
  facet_wrap(~ strata, scales = "free") +
  ggtitle("Number of trees per community along slope gradient.")
left_join(trees, filter(PFT, trait == "WD")) %>% 
  filter(strata == "U") %>% 
  filter(!is.na(traitvalue)) %>% 
  mutate(WDclass = cut(traitvalue, 
                       breaks = c(0, 0.5, 0.7, 1), 
                       labels = c("low", "mid", "high"))) %>% 
  group_by(WDclass, community) %>% 
  summarise(Abundance = sum(Abundance)) %>% 
  left_join(R) %>% 
  ggplot(aes(Slope, Abundance)) +
  geom_point() +
  facet_wrap(~ WDclass, scales = "free", nrow = 2) +
  ggtitle("Number of trees per community along slope gradient",
          "depending on WD category (low < 0.5 and high > 0.7).")
```

1. Sylvain Schmitt
    - Corresponding author
    - sylvain.m.schmitt@gmail.com
    - orcid.org/0000-0001-7759-7106
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
    - UMR BIOGECO, INRA - Université de Bordeaux, 69 route d’Arcachon, F-33612 Cestas cedex France
1. Valérie Raevel
    - vraevel@gmail.com
    - orcid.org/0000-0002-9631-114X
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
1. Maxime Réjou-Méchain
    - maxime.rejou@ird.fr
    - orcid.org/0000-0003-2824-267X
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
    - AMAP, IRD, CNRS, CIRAD, INRA, Univ Montpellier, Montpellier, France
1. Narayanan Ayyappan
    - ayyappan.n@ifpindia.org
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
1. Natesan Balachandran
    - balachandran.n@ifpindia.org
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
1. N. Barathan
    - barathan.n@ifpindia.org
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
1. Chandra Shekhar Jha
    - chandra.s.jha@gmail.com
    - National Remote Sensing Centre, Indian Space Research Organisation, Hyderabad 500 037, India 
1. François Munoz
    - fmunoz@univ-grenoble-alpes.fr
    - orcid.org/0000-0001-8776-4705
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
    - Laboratoire d'Écologie Alpine, Univ. Grenoble Alpes, FR-38000 Grenoble, France

@@

# Changes

## 13 octobre 2020

* I changed the title.
* I removed the bidirectional stepwise selection of predictors as it was unstable.
* I used bayesian inference for the linear regression (minor).
* I replaced table 2 by a figure of posteriors distribution and significance (seemed clearer).
* I removed the ideas of LES and WES between strata. That's unclear and difficult to defend. WD model in understorey showed a $R^2=0.3$, which is huge and removed the idea that the WES mainly affects canopy. Similarly, SLA and LDMC models in canopy had low $R^2$ but showed significant effects of wetness, curvature and slope.
* I kept Maxime’s hypothesis for canopy height effect in understory. I don’t have a better hypothesis. I am therefore to keep it calling for further testing. **Maybe we should remove canopy height to simplify the message?**
* I am not a fan of the conclusion. I kept it for the moment but I didn’t feel it brought much more than the take home message; So I’m hesitating to simply remove it.

@@

1. Topographic heterogeneity determines changing soil conditions, while the height of the canopy modulates the microclimate in the forest. But the environmental conditions encountered by trees also change across the vertical strata of tropical forests. However, the joint influence of topography and canopy height with vertical strata on tree functional composition remain little known. We hypothesized different leaf and wood functional strategies between canopy and understory guilds of trees, and different responses of the guilds to topographic heterogeneity and canopy height, depending on resource availability, hydraulic and biomechanical constraints.
1. We measured leaf and wood traits over 89 species representing 99% of all individuals in a 10-ha rainforest plot located in the Western Ghats, South India. The plot was subdivided into tree communities, each composed of guilds of canopy and understory trees. We combined tree census data, LiDAR-derived topographic and canopy metrics, and functional traits to assess the variation of functional composition in understory and canopy guilds, and to address the effects of abiotic and biotic environments on these compositions.
1. The variation of functional traits across sampled tree species followed locally decoupled wood and leaf economics spectra. Although functional space did not differ overall between canopy and understory tree species, environmental filtering shaped different functional composition of canopy and understory guilds in tree communities. Furthermore, the functional composition of the canopy and understory guilds responded differently across communities to environmental gradients related to soil resource, light availability and aspect.
1. *Synthesis.* These results show that both fine-scale environmental heterogeneity and vertical structure shape the composition of hyperdiverse rainforest tree communities. Guild-specific responses to biotic and abiotic environmental filtering underline that vertical stratification must be acknowledged to better understand and forecast forest dynamics.

@@

Rainforest; Western Ghats; Economics spectra; Environmental filtering, Vertical strata

@@

Understanding the drivers of tree composition in hyperdiverse tropical forests is a long-standing issue in ecology [@Cody1976]. 
Environmental variation influences the composition of tree community at regional [@Rejou-Mechain2008; @Tuomisto2003] 
and local scale [@Ferry2007; @gunatilleke_specieshabitat_2006]. 
Topography is known to filter different ecological strategies [@kraft_functional_2008; @Schmitt2020]
due to its impact on water and nutrient availability [@Blanchard2019; @ferry2010higher].
In addition, canopy height modulates forest microclimates due to topography [@Jucker2018].
But, tree species  occupy different vertical strata at adult stage, and can thus be classified into different vertical guilds (understory and canopy),
experimenting contrasted environmental conditions. 
Indeed, light availability decreases downward across vertical forest strata [@chazdon1984],
while hydraulic and mechanical constraints increase upward [@Koch2004], 
so that different survival and growth strategies are filtered across strata [@Raevel2018; @Stark2015; @Uriarte2004a]. 
While functional strategies of tree species have been documented independently across topography, canopy height and vertical guilds,
little is known on their interactions,
particularly the effect of topography and canopy height on species assemblages between vertical guilds has not been addressed [@wright2002plant].

Functional traits express fundamental trade-offs determining species occurrences and co-occurrences along environmental gradients [@Wright2002], 
contributing thus to test for community assembly processes. 
Functional traits have been defined as phenotypic traits that impact fitness through their effect on individual performance, which is defined as the ability to recruit, grow, survive, and reproduce in a given environment [@violle_let_2007].
Functional traits covary among species [@diaz_global_2015] and communities [@Bruelheide2018
along distinct economics spectra of leaf [@wright_worldwide_2004; @osnas_global_2013; but see @lloyd_photosynthetically_2013] and wood [@chave_towards_2009].
The leaf economics spectrum expresses a trade-off between resource acquisition by photosynthesis and investment in leaf defense and durability [@Evans2001a; @hodgson_is_2011; but see @Gibert2016].
Trait values along the leaf economics spectrum are expected to change along soil and light availability gradients. 
The wood economics spectrum opposes slow growing, drought-resistant species to fast growing, drought-intolerant species [@chave_towards_2009].
Previous studies showed decoupled trait variations along the leaf and wood economics spectra for tropical trees at large scale in the Neotropics [@Baraloto2010; @fortunel_leaf_2012]. 
Nevertheless, @reich_world-wide_2014
proposed a unique plant economics spectrum simply opposing fast and slow growing species as observed in subarctic flora [@freschet_evidence_2010].
The concept of plant economics spectrum may be highly relevant for understanding the variation of strategy between vertical guilds as understory species are generally characterized by a low potential growth rate associated to a conservative strategy of resource use, while canopy species tend to have a higher potential growth rate associated to a faster resource acquisition [@Herault2011].

In the context of seasonal rainforests, we hypothesized different trait-environment relationships across vertical guilds, 
with varying strategies along the fundamental leaf and wood economics spectra. 
More specifically, because understory trees are strongly light-limited and are less subject to hydraulic and mechanical constraints than canopy trees, 
we expect that  understory and canopy species tend to maximize their difference along the leaf and wood economic spectra, respectively. 
More generally, we expect that understory species and assemblages have more conservative strategies than canopy ones and that the lower water and nutrient availability found on hilltops [@ferry2010higher] 
entail stronger conservative resource-use strategy [@Cunningham2007; @kraft_functional_2008; @Mendez-Toribio2017; @Schmitt2020]. 
Previous works have shown different response of vertical strata to environmental variation at ground level, with increased edaphic constraint on understory compared to canopy  for Columbian drained uplands [@duque2002different].

Here, we assessed the variation in functional composition of understory and canopy guilds within tropical tree communities, 
and examined the effects of topography and canopy height on these variations. 
We measured leaf and wood traits of 89 species representing 99% of all individuals in a 10-ha forest plot located in the Western Ghats, South India, a region considered as a global biodiversity hotspot [@myers_biodiversity_2000].
The plot encompasses a broad variation of topography [@Antin2013] and vertical structure [@Pascal1988].
Because the region experiences a long dry season period and is under the influence of intense monsoon during rainy seasons with strong wind disturbances, 
we expected hydraulic and mechanical constraints to play an important role in tree community assembly.

@@

@@

Using the classification of @Pascal1988, we assigned a potential vertical stratum to each tree species corresponding to the vertical stratum reached by most adult trees.
The strata included four categories: emergent above 39 m, canopy between 25 and 39 m, sub-canopy between 15 and 25 m, and understory below 15 m. 
Because our study site is characterized by a steep topography (Fig. \@ref(fig:Uppangala)), 
sub-canopy species can often reach the canopy in the Uppangala plot. 
We thereby defined two broader categories, namely, canopy guild (C) including emergent, 
canopy and sub-canopy species and understory guild (U) composed of understory trees.

We collected trait data for 89 over the 111 (80%) species, representing 99% of tree individuals in the plot (54 canopy and 35 understory species). We sampled five individuals per species for the 32 most abundant species (>80% of individuals), and one to five individuals for the remaining species (3 individuals on average), depending on leaf accessibility and species abundance. For each individual, we collected at 4 to 12 m height five mature leaves and one branch sample with a diameter between 1 and 2 cm, following the recommendation of @Perez-Harguindeguy2013. Whenever possible, individuals and leaves were selected in similar light conditions (Dawkins index of 2; [@Dawkins1958]). Leaf and branch samples were kept in humidified ziplock bags filled with enriched $CO_2$ air in dark until measurement within the next 24 hours. For 11 species (1% of individuals in Uppangala plot), branch samples were too small (< to 1 cm in diameter) to allow measuring branch wood density.

@@

if(knitr:::is_latex_output()){
  t %>% 
    ungroup() %>% 
    mutate(Trait = c('Leaf area', 'Leaf dry\nmatter content', 'Leaf\nthickness', 'Specific\nleaf area', 'Wood\ndensity'),
    Strategy = c('Leaf defense\nand investment',
                 'Leaf resource\ncapture vs. investment',
                 'Leaf defense\nand investment',
                 'Leaf resource\nacquisition vs. defense',
                 'Stem transport,\nstructure and defense')) %>% 
    mutate_all(linebreak) %>% 
    kable(caption = 'Functional traits measured. Traits are presented with their abbreviation, unit, related species strategy according to @baraloto_decoupled_2010, range of values, intra-individual, intra-specific and inter-specific coefficients of variation (CV).',
          digits = 2, format.args = list(big.mark = " "), format = 'pandoc', escape = F,
          col.names = c("Trait", "Abbreviation", "Unit", "Strategy", "Range", 
                        "\\makecell[l]{Intra-\\\\individuals}", 
                        "\\makecell[l]{Intra-\\\\species}", 
                        "\\makecell[l]{Inter-\\\\species}")) %>%
    kable_styling(full_width = T)
} else {  
  t %>% 
  kable(digits = 2, format.args = list(big.mark = " "), format = 'pandoc', escape = F,
                  col.names = c("Trait", "Abbreviation", "Unit", "Strategy", "Range", 
                        "Intra-individuals", "Intra-species", "Inter-species"),
        caption = 'Functional traits measured. Traits are presented with their abbreviation, unit, related species strategy according to @baraloto_decoupled_2010, range of values, intra-individual, intra-specific and inter-specific coefficients of variation (CV).') %>% 
  kable_styling(full_width = T)
}
```

@@

We subsampled tree communities in $30~m*30~m$ subplots (n=110), in order to characterize the functional composition of vertical guilds within each subplot. Each subplot contained 28 individuals and 10 species on average, with less species in understory (6 in average) than in canopy (13 in average) guilds. We thus calculated the weighted mean trait value per guild (GWM, where G stands for ‘guild’; [@Raevel2018]) for each trait in each subplot:

@@

Functional traits data have been submitted to TRY initiative [@Kattge2019] under the name UppangalaTraits. 
Environmental variables and spatial positions of individuals were extracted from the Uppangala Station database.
Access to LiDAR data is restricted by the Indian government and should be the object of a convention (contact: chandra.s.jha@gmail.com).
